---
title: "Week 7 homework Rmarkdown"
author: "**Statistical Analysis**"
date: "*Week 7*"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true    
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dataset 1 analysis

```{r, echo = FALSE, results = 'hide', message = FALSE}
library(tidyverse)
library(vroom)
library(usethis)
library(devtools)
library(lubridate)
library(multcomp)
library(scales)
library(glmmTMB)
library(DHARMa)
library(fitdistrplus)
library(MuMIn)
library(MASS)
library(viridisLite)
library(viridis)
library(ggplot2)

```

Read in dataset 1 from github
```{r, eval = FALSE}
dataset_1 <- vroom("https://raw.githubusercontent.com/eh17393/bioinformatics/main/Bioinformatics_data-master/Workshop%206/data%201.csv")

```
```{r, echo = FALSE, results = 'hide', message = FALSE}
dataset_1 <- vroom("https://raw.githubusercontent.com/eh17393/bioinformatics/main/Bioinformatics_data-master/Workshop%206/data%201.csv")
```
```{r}
dataset_1
```

## Visualise dataset 1 with x1
```{r}
ggplot(dataset_1, aes(x = y, y = x1)) +
  geom_point() +
  geom_smooth(method = "lm")
```
No significant correlation

## Visualise dataset 1 with x2
```{r}
ggplot(dataset_1, aes(x = y, y = x2)) +
  geom_point() +
  geom_smooth(method = "lm")
```
Potentially a significant positive correlation 

Fit a glm with gaussian distribution and plot
```{r}
mod1_d1x2 <- glm(x2 ~ y, data = dataset_1, family = "gaussian")
plot(mod1_d1x2)
```
Summarise mod 1
```{r}
summary(mod1_d1x2)
```
Significant intercept however plots are not perfect therefore test with other models with gaussian distribution

Fit a glm with a gaussian distribution with a log link
```{r}
mod2_d1x2 <- glm(x2 ~ y,
                   data = dataset_1,
                   family = gaussian(link = "log"))
plot(mod2_d1x2)
```
Fit glm with a guassian model with an inverse link
```{r}
mod3_d1x2 <- glm(x2 ~ y,
                   data = dataset_1,
                   family = gaussian(link = "inverse"))
plot(mod3_d1x2)
```
Compare models with AIC and rank
```{r}
library(gamlr)
##compare the models 
AIC_mods <- data.frame(model = c("mod1_d1x2", "mod2_d1x2", "mod3_d1x2"),
                       AICc = c(AICc(mod1_d1x2), AICc(mod2_d1x2), AICc(mod3_d1x2)))

##rank the models
AIC_mods[order(AIC_mods$AICc),]
```
Mod 1 the best therefore gaussian error distribution with significant x2 

## Visualise dataset 1 with x3
```{r}
ggplot(dataset_1, aes(x = y, y = x3)) +
  geom_point() +
  geom_smooth(method = "lm")
```
Potentially a significant correlation with y 

Fit a glm with gaussian distribution and plot
```{r}
mod1_d1x3 <- glm(x3 ~ y, data = dataset_1, family = "gaussian")
plot(mod1_d1x3)
```
Summarise mod 1 
```{r}
summary(mod1_d1x3)
```
Significant correlation with intercept therefore x3 is significant with gaussian error distribution

## Visualise dataset 1 with x4
```{r}
ggplot(dataset_1, aes(x = y, y = x4)) +
  geom_point() +
  geom_smooth(method = "lm")
```
No significant correlation 


# Dataset 2 analysis

Read in dataset 2 from github
```{r, eval = FALSE}
dataset_2 <- vroom("https://raw.githubusercontent.com/eh17393/bioinformatics/main/Bioinformatics_data-master/Workshop%206/data%202.csv")
```
```{r, echo = FALSE, results = 'hide', message = FALSE}
dataset_2 <- vroom("https://raw.githubusercontent.com/eh17393/bioinformatics/main/Bioinformatics_data-master/Workshop%206/data%202.csv")
```
```{r}
dataset_2
```
## Visualise dataset_2 with x1
```{r}
ggplot(dataset_2, aes(x = y, y = x1)) +
  geom_point() +
  geom_smooth(method = "lm")
```
Fit a glm with gaussian distribution and plot 
```{r}
mod1_d2x1 <- glm(x1 ~ y, data = dataset_2, family = "gaussian")
plot(mod1_d2x1)
```
Normal Q-Q plot doesn't fit very well therefore try with other glms

Fit glm with poisson distribution
```{r}
mod2_d2x1 <- glm(x1 ~ y, data = dataset_2, family = "poisson")
plot(mod2_d2x1)
```
Fit a glm with a gaussian distribution with a log link and plot
```{r}
mod3_d2x1 <- glm(x1 ~ y,
               data = dataset_2,
               family = gaussian(link = "log"))
plot(mod3_d2x1)
```

Fit glm with a guassian model with an inverse link
```{r}
mod4_d2x1 <- glm(x1 ~ y,
               data = dataset_2,
               family = gaussian(link = "inverse"))
plot(mod4_d2x1)
```

Compare and rank models with AIC
```{r}
##compare the models 
AIC_modsd2 <- data.frame(model = c("mod1_d2x1", "mod2_d2x1", "mod3_d2x1", "mod4_d2x1"),
                       AICc = c(AICc(mod1_d2x1), AICc(mod2_d2x1), AICc(mod3_d2x1), AICc(mod4_d2x1)))

##rank the models
AIC_modsd2[order(AIC_mods$AICc),]
```
Mod 1 is the best fit showing x1 is significant with a gaussian error distribution


## Visualise dataset_2 with x2 
```{r}
ggplot(dataset_2, aes(x = y, y = x2)) +
  geom_point() +
  geom_smooth(method = "lm")
```

## Visualise dataset_2 with x3
```{r}
ggplot(dataset_2, aes(x = y, y = x3)) +
  geom_point() +
  geom_smooth(method = "lm")
```




# Dataset 4 analysis

Read in dataset 4 from github
```{r, eval = FALSE}
dataset_4 <- vroom("https://raw.githubusercontent.com/eh17393/bioinformatics/main/Bioinformatics_data-master/Workshop%206/data%204.csv")
```
```{r, echo = FALSE, results = 'hide', message = FALSE}
dataset_4 <- vroom("https://raw.githubusercontent.com/eh17393/bioinformatics/main/Bioinformatics_data-master/Workshop%206/data%204.csv")
```
```{r}
dataset_4
```
## Visualise dataset 4 with x1
```{r}
ggplot(dataset_4, aes(x = y, y = x1)) +
  geom_point() +
  geom_smooth(method = "lm")
```
Shows a potentially strong positive correlation with y

Fit a glm with gaussian distribution, plot and summarise
```{r}
mod1_d4x1 <- glm(x1 ~ y, data = dataset_4, family = "gaussian")
plot(mod1_d4x1)
```
```{r}
summary(mod1_d4x1)
```
Fit glm with poisson distribution, plot and summarise
```{r}
mod2_d4x1 <- glm(x1 ~ y, data = dataset_4, family = "poisson")
plot(mod2_d4x1)
```
The residuals vs leverage graph is better than gaussian

```{r}
summary(mod2_d4x1)
```
Not significant therefore log x1 and summarise
```{r}
mod2log_d4x1 <- glm(log10(x1) ~ y, data = dataset_4, family = "poisson")
summary(mod2log_d4x1)
```
x1 is significant with poisson error distribution and log link function

## Visualise dataset_4 with x2
```{r}
ggplot(dataset_4, aes(x = y, y = x2)) +
  geom_point() +
  geom_smooth(method = "lm")
```
A clear positive correlation with y 

Fit glm with log link and poisson distribution and summarise
```{r}
mod2log_d4x2 <- glm(log10(x2) ~ y, data = dataset_4, family = "poisson")
summary(mod2log_d4x2)
```
x2 is significant with poisson error distribution and log link function

## Visualise dataset_4 with x3
```{r}
ggplot(dataset_4, aes(x = y, y = x3)) +
  geom_point() +
  geom_smooth(method = "lm")
```
No clear correlation









