---
title: "Week 6 Homework Rmarkdown"
author: "**Statistical Analysis**"
date: "*Week 6*"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true    
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tokyo Olympics 2021 Medal Data
```{r, echo = FALSE, results = 'hide', message = FALSE}
library(tidyverse)
library(vroom)
library(usethis)
library(devtools)
library(lubridate)
library(multcomp)
```


Read in the Tokyo 2021 olympic medal data from github
```{r, eval = FALSE}
medal_data <- vroom("https://raw.githubusercontent.com/chrit88/Bioinformatics_data/master/Workshop%205/Tokyo%202021%20medals.csv")
```
```{r, echo = FALSE, results = 'hide', message = FALSE}
medal_data <- vroom("https://raw.githubusercontent.com/chrit88/Bioinformatics_data/master/Workshop%205/Tokyo%202021%20medals.csv")
```

```{r}
medal_data
```

Add GDP column for all countries by calling `library(wbstats)`
```{r, echo = FALSE, results = 'hide', message = FALSE}
library(wbstats)
```
Extract GDP data for all countries for 2020 (no 2021 data)
```{r}
all_GDP_data <- wb_data(indicator = "NY.GDP.MKTP.CD",
                        start_date = 2020,
                        end_date = 2020)
all_GDP_data
```
Rename all_GDP_data column 5 to GDP
```{r}
names(all_GDP_data)[5] <- "GDP"
all_GDP_data
```
Ensure all countries are identified as the correct country before merging with world data by calling `library(countrycode)`
```{r,  echo = FALSE, results = 'hide', message = FALSE}
library(countrycode) 
```
Add a column to medal data containing the country code
```{r}
medal_data$code <- countrycode(medal_data$Country, 
                               origin = "country.name",
                               destination = "iso3c")
medal_data
```
Look at head of new dataset
```{r}
head(medal_data, 3)
```
Correct China
```{r}
medal_data$code[2] <- "CHN"
head(medal_data, 3)
```
Select just iso3c and GDP column
Use dplyr::select as conflict with MASS package
```{r}
head(all_GDP_data %>% dplyr::select( iso3c, GDP ))
```
Join two data sets 
```{r}
medal_GDP <- left_join(medal_data, 
                         all_GDP_data %>% dplyr::select(iso3c, GDP),
                         by = c("code" = "iso3c"))
medal_GDP
```
Drop rows with NA values
```{r}
medal_GDP <- medal_GDP %>% drop_na()
```
As gold medal frequency essentially determines table ranking, plot gold medals won against country GDP
```{r}
p1 <- ggplot(medal_GDP, aes(x = GDP,
                             y = Gold)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  ylab("Gold Medals Won") +
  xlab("GDP ($)")

p1 <- p1 + geom_smooth(method="loess")

p1 <- p1 + ggtitle("Country GDP versus Golds won at Tokyo 2021")
p1
```
Fit a glm()
```{r}
mod_gold_1 <- glm(Gold ~ GDP,
                  data = medal_GDP,
                  family = "gaussian")
mod_gold_1
```
Plot diagnostic plots
```{r}
plot(mod_gold_1)
```
Summarise the model outputs 
```{r}
summary(mod_gold_1)
```
Significant impact of GDP on gold medals won: higher GDP correlates with higher gold medal count.


Fit a glm with a poisson distribution to see if better model fit.
```{r}
mod_gold_2 <- glm(Gold ~ GDP,
                  data = medal_GDP,
                  family = "poisson")
plot(mod_gold_2)
```

Not good fit as normal Q-Q plot very ill-fitting plus points are outside 1 lines on residuals vs leverage graph

# Iris flower data

Load the iris data set
```{r}
data("iris")
```
plot the Sepal.Width data by Sepal.Length and Species 
```{r}
ggplot( iris, aes( x = Petal.Length, y = Sepal.Width)) +
  geom_jitter(aes(col = Species)) +
  theme_bw()+
  ggtitle("How petal length impacts sepal width broken down by species")
```
Fit a glm()
```{r}
mod_petal <- glm(Sepal.Width ~ Petal.Length*Species,
                data = iris,
                family = "gaussian")
mod_petal
```
Plot model
```{r}
plot(mod_petal)
```

Summarise model outputs
```{r}
summary(mod_petal)
```
To achieve a better fit transform response variable Sepal.Width with log10()
```{r}
log_sepal_width <- log(iris$Sepal.Width)
```

Add log_sepal_width column to dataframe
```{r}
iris <- iris %>% mutate(log_sepal_width)
```
Plot the log_sepal.width data by sepal.length and species 
```{r}
ggplot( iris, aes( x = Petal.Length, y = log_sepal_width)) +
  geom_jitter(aes(col = Species)) +
  theme_bw()+
  ggtitle("How petal length impacts the log10 of sepal width broken down by species")

```
Fit a glm()
```{r}
mod_log <- glm(log_sepal_width ~ Petal.Length*Species,
                 data = iris,
                 family = "gaussian")
mod_log
```
Plot model
```{r}
plot(mod_log)
```

Summarise model outputs
```{r}
summary(mod_log)
```
More significant than previous glm showing versicolor and and virginica affected by sepal width however no overall effect of petal.length.









